Optifcal Observations of an Extraordinarily Young Type Ia Supernova iPTF16abc

### 2017-01-02 doc/ms.tex
v0.1

\title{iPTF16abc}

@@

A transient, designated as iPTF16abc, was discovered by the
intermediate Palomar Transient Factory on 2016 April
$3.36$\footnote{all times in this paper are in UTC.} at
$\textrm{R.A.}=13^h34^m45.49^s$, $\textrm{Dec.}=+13^d51^m14.3^s$
(J2000) with a $g$-band magnitude of $21.31\pm0.27$
\citep{2016PASP..128k4502C,2016ATel.8907....1M}. The transient is
spatially coincident with a tidal tail of the galaxy NGC\,5221 at
100\,Mpc. No activity was detected at the same location down to
$g=22.1$\,mag on April $2.42$. Our spectroscopic follow-up compaign
classified iPTF16abc as a normal SN Ia \citep{2016ATel.8909....1C}.

This paper is organized as follows:

@@

@@

After discovery, we also utilized the rainbow camera of the SED
Machine (\ycao{REF}) mounted on the 60-inch telescope at Palomar
Observatory (P60) to carry out photometric observations in $g$, $r$
and $i$ filters. The image differencing against the archival SDSS
images and forced PSF photometry on the subtracted images were
performed by the Fremling Automated Pipeline
\citep{2016A&A...593A..68F}. The photometry is also calibrated to the
SDSS catalog.

@@

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.9\textwidth]{spec_sequence.pdf}
  \caption{Low-resolution spectra of iPTF16abc are shown in the
    chronical sequence. The phases in units of days are labeled next
    to corresponding spectra. Telluric absorption bands are grayed
    out.}
  \label{fig:spec_seq}
\end{figure*}

@@

The foreground Galactic extinctioin along the directionof iPTF16abc ha
$E(B-V)=0.0279\,\textrm{mag}$ \citep{2011ApJ...737..103S}.

The UVES spectrum of iPTF16abc shows narrow double absorption features
from \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D lines (Figure
\ref{fig:narrow_features}), indicating two sources of absorption along
the line of sight. Fitting two Gaussian kernels to each of
the \ion{Na}{1} doublet simultaneously leads to redshifts of
$0.02313820\pm0.00000032$ and $0.02322408\pm0.00000033$. The total
equivalent widths of the \ion{Na}{1}\,D1 and D2 lines are
$0.595\pm0.009\,\textrm{\AA}$ and $0.609\pm0.008\,\textrm{\AA}$.
Using the empirical relation in \citet{2012MNRAS.426.1465P}, we derive
a total extinction of $E(B-V)=0.361\pm0.025\textrm{mag}$.

The X-shooter spectra of iPTF16abc also shows absorption features at
\ion{K}{1}\,7665\,\AA and 7699\,\AA. However, the resolution is not
high enough for the spectra to resolve the two components seen in the
profiles of \ion{Ca}{2} and \ion{Na}{1} in the UVES spectrum. The
X-shooter spectra do not show the diffusive interstellar band at
5780\,\AA as well.  Albeit some arguments (e.g.,
\citealt{2013ApJ...779...38P}), the extinction derived from
\ion{Na}{1}\,D absorption provides the best estimate in the case of
iPTF16abc.

@@

The \ion{Na}{1}\,D doublet are resolved in multiple spectra spanning
from pre-peak to post-peak phases. Despite the instrumental widening of
different instrument configurations, we do not detect obvious variation
in the profiles of the doublet.

@@

We run Supernova Identification (SNID; \citealt{2007ApJ...666.1024B})
on the low-resolution spectrum of iPTF16abc at $+18.8$ and found best
matches to normal SNe Ia. Separately the characteristic features of a
SN Ia, such as \ion{Si}{2}, \ion{S}{2}, are obviously seen in the
spectra of iPTF16abc.

The optical light curves obtained by P48 and P60, after correction for
extinction, are fitted by the SALT2 template
\citep{2007A&A...466...11G} with the \texttt{sncosmo} Python
module\footnote{The \texttt{sncosmo} module is available at
  \url{https://sncosmo.readthedocs.io/en/v1.4.x/}.}. The fitting
parameters are $t_{max}=57499.65\pm0.02$, $x_0=0.0275\pm0.0002$,
$x_1=1.200\pm0.043$, and $c=-0.3353\pm0.0054$. The best-fit model also
gives an unreddened apparent peak magnitude of
$m^*_{B}=14.4\,\textrm{mag}$ in the SN rest frame.

For convenicne, in the following anlaysis and discussion, we define
$t=MJD - 57499.65\,\textrm{days}$, i.e., the time with respect to the
best-fit $t_{max}$.

@@

After establishing iPTF16abc as a normal SN ia, we use the latest
calibration \citep{2014A&A...568A..22B} of the Phillips relation
\citep{1993ApJ...413L.105P} with $m^*_{B}$, $x_1$ and $c$ to derive a
distance modulus $\mu=34.66\pm0.03\,\textrm{mag}$, provided that the
host galaxy of iPTF16abc has a stellar mass less than $10^{10}\sm$.

The location of iPTF16abc is spatially coincident with a tidal tail of
galaxy NGC\,5221. \citet{2007A&A...465...71T} observed this galaxy in
the near-infrared and derived a distance modulus of $35.0\pm0.4\,\textrm{mag}$
from the Tully-Fisher relation. This distance modulus is consistent
with that of iPTF16abc.

Separately, \cite{1998A&AS..130..333T} observed
the 21-cm line in this galaxy and measured a redshift of $0.0233303\pm0.000027$.
The two components in the \ion{Na}{1}\,D have a relative velocity of
$-57.6\pm8.1\,\textrm{km}\,\textrm{s}^{-1}$ and
$-31.8\pm8.1\,\textrm{km}\,\textrm{s}^{-1}$, suggesting that both absorption
resources are probaly located on the tidal tail of NGC\,5221.

\section{First Light And Early Rise}
\label{sec:early_phases}

In this section, we analyze the early light curve and spectra of iPTF16abc.

@@

In order to estimate the start time of the observed light curve, denoted as $t_0$,
we fit a power-law model
\begin{equation}
  \label{eq:broken_power_law}
  f(t) \left\{
    \begin{array}{ll}
      = 0,\ \textrm{when}\ t<t_0 \\
      \propto (t-t_0)^{\alpha},\ \textrm{when}\ t>t_0
    \end{array}
  \right.
\end{equation}
to the early-phase \textit{g}-band light curve. We experiment the
fitting procedure with different time windows and find that the light
curve between $t=-18\,\textrm{days}$ and $t=-14\,\textrm{days}$
increases approximately linearly. The best fit model of
$\alpha = 0.97$ and $t_0=-18.47$ is shown in the left panel of
Figure \ref{fig:early_lc_fit} and the join probability distribution of
$\alpha$ and $t_0$ is in the right panel of the same figure.  With the
best-fit $t_0$, our first observation was made only
$0.18\,\textrm{day}$ after the rise of the light curve.

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.45\textwidth]{early_lc.pdf}
  \includegraphics[width=0.45\textwidth]{rise_time_power_law_index.pdf}
  \caption{Broken Power low fitting to the early $g$-band light
    curve. \textit{Left:} the best-fit model of $\alpha=0.94$ and
    $t_0=-18.47\,\textrm{days}$ is illustrated against the
    data. \textit{Right:} the joint distribution of $t_0$ and
    $\alpha$. The cross marker denotes the best-fit parameters. The
    inner and outer contours represent the $68\%$ and $99.7\%$
    confidence levels.}
  \label{fig:early_lc_fit}
\end{figure*}

The \textit{g}-band light curve beyond $t=-14\,\textrm{days}$ rises
significantly faster than the best-fit model above, indicating a
larger value of $\alpha$. In fact, if we fit the light curve between
$t=-14$ and $t=-8$\,days to another power law, we find the best fit at
a power-law index of $1.40$ (Figure \ref{fig:against_sn2012cg}).

Although the early rise behavior of iPTF16abc is distinct from that of
SN2011fe \citep{2011Natur.480..344N} which follows the $t^2$ Arnett
law \citep{1982ApJ...253..785A}, we have seen rise behavior similar to
iPTF16abc (i.e., a steep initial rise followed by a more gradual and
steadier rise) in a few nearby events that were observed at extremely
young phases, such as, SN2013dy \citep{2013ApJ...778L..15Z}, SN2014J
\citep{2014ApJ...783L..24Z}, and ASASSN-14lp
\citep{2016ApJ...826..144S}.

We also note statistical studies on rise
behavior of SNe Ia (e.g., \citealt{2010ApJ...712..350H, 2011MNRAS.416.2607G,
  2012ApJ...745...44G, 2015MNRAS.446.3895F}) which suggest a power-law
index $\gtrapprox2$. However, these studies are not sensitive to SNe Ia
at extremely early phases, so the initial quasi-linear rise of iPTF16abc
is not comparable to their conclusions. On the other hand, the subsequent
$t^{1.40}$ rise of iPTF16abc is roughly consistent with these samples.

@@

As pointed out by \citet{2014ApJ...784...85P}, if a SN light curve is
purely powered by radioactive decay of $^{56}$Ni, the SN may
experience a dark period between the explosion time of the SN
$t_{exp}$ and the rise time of its light curve $t_0$, because it takes
time for radioactive energy to diffuse from where $^{56}$Ni is
deposited in the ejecta and the SN photosphere. Therefore,
\citet{2014ApJ...784...85P} suggests to use the early-phase spectra to
measure line velocities and then estimate $t_{exp}$ by assuming that
$v\propto (t-t_{exp})^{-0.22}$.

We follow the same procedure and
measure the velocities of the \ion{Si}{2}\,6355 line. Because we do
not see signals of multiple velocity components in the
\ion{Si}{2}\,6355 line, and because the \ion{Si}{2}\,6355 line is
partially blended with the \ion{C}{1}\,6580 line, we fit two gaussian
kernels and a linear term simultaneously to spectra between about
6100\,\AA and about 6600\,\AA in the rest frame. Then the velocity is
measured at the minimum of the Gaussian kernel that corresponds to the
\ion{Si}{2} feature.

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.45\textwidth]{exp_date_chi2.pdf}
  \includegraphics[width=0.45\textwidth]{SiIIVelocity.pdf}
  \caption{Constraints on $t_{exp}$ from fitting the velocity
    evolution of \ion{Si}{2}. \textit{Left panel:} the dashed, solid
    and dash-dotted curves show $\chi^2$ for fitting power laws with
    indices $-0.20$, $-0.22$ and $-0.24$, respectively. The red
    vertical line and the orange region indicate $t_0$ and its
    3-$\sigma$ confidence interval from Section
    \ref{sec:lc_fit}. \textit{Right panel:} Observed \ion{Si}{2}\,6355
    velocities with the best-fit power-law velocity with an index of
    $-0.22$.}
  \label{fig:velocity_t_exp}
\end{figure*}

The expansion velocity fit (Figure \ref{fig:velocity_t_exp}) shows
that the best-fit explosion time $t_{exp}=-17.95\,\textrm{days}$ with
a 3-$\sigma$ confidence interval between $-17.25\,\textrm{days}$ and
$-18.65\,\textrm{days}$.

We further test our assumption $v\propto (t-t_{exp})^{-0.22}$ by
altering the power-law index to $-0.20$ and $-0.24$ (Figure
\ref{fig:velocity_t_exp}).  The best-fit value of $t_0$ varies within
the 3-$\sigma$ confidence interval and the $\chi^2$ value of the
best-fit model varies very little. This test verifies the robustness
of the assumed power-law index of $-0.22$.

Comparing $t_{exp}$ and $t_0$ (Section \ref{sec:lc_fit}), we conclude
that the dark phase of the SN, if exists, is very brief. In the
following analysis, it is reasonable to assume $t_{exp}\simeq t_0$.

\subsection{Light Curve Energy Resources}
\label{sec:lc_energy}

The early portion of a SN light curve may have multiple resources: SN
shockbreakout, SN-companion collision, and radioactive activity. Since
each provides interesting constraints on the progenitor properties, we
explore all the three posibilities.

First, the shock breakout of a SN Ia lasts for a fraction of a second
due to the small size of the exploding star. However, the subsequent
cooling phase may last longer (e.g., \citealt{2010ApJ...708..598P}).
Following the analysis of SN2011fe in \citet{2012ApJ...744L..17B}, we
compare the early-phase \textit{g}-band light curve of iPTF16abc with
two cooling models \citep{2011ApJ...728...63R, 2010ApJ...708..598P}
and reach a not very constraining conclusion that the radius of the
progenitor star of iPTF16abc should be $<1\sr$. In fact, the cooling
emission is negligible at the time of the first detection of iPTF16abc.

Second, if iPTF16abc is born in a single-degenerate channel, with a
chance of $\lesssim10\%$, we expect to see emission produced by SN
ejecta slamming into the companion. In the following we argue from a
statistical \ycao{TBW}

In the case of SN2012cg, \citet{2016ApJ...820...92M} fitted power-law
models to its light curves between $t=-14$ days and $t=-8$ days and
extrapolated backwards to illustrate excess fluxes at $t<-14$
days. The authors further attempted to explain the excess fluxes as
collisional signatures between the SN ejecta and a non-degenerate
companion \citep{2010ApJ...708.1025K}. If we perform the same practice
to the iPTF16abc data, then we find similar excess fluxes at earlier
phases (Figure \ref{fig:against_sn2012cg}). In fact, for any SN whose
light curve can be approximated by a broken-power law, i.e.,
\begin{equation}
  \label{eq:broken-pow}
  f(t) = t^{\alpha}(1 + t^{s(\beta - \alpha)})^{1/s}\ ,
\end{equation}
one can always fit the relative late part of the light curve with a
power law model and find an excess in the early phases.

\begin{figure}[htb]
  \centering
  \includegraphics[width=0.45\textwidth]{another_early_lc.pdf}
  \caption{The early-phase \textit{g}-band light curve of iPTF16abc is
    compared against the best-fit power-law model to the data segment
    between $t=-14$ and $t=-8$ days. Note the ``excess'' between the
    data and the model during $t=-18$ and $t=-15$ days}
  \label{fig:against_sn2012cg}
\end{figure}

As a result of the above analysis, we conclude that the early light
curve of iPTF16abc is probably powered dominantly by radioactive decay
of $^{56}$Ni.  According to calculations in
\citet{2016ApJ...826...96P}, when synthesized $^{56}$Ni is mixed
sufficiently in the ejecta, the dark period is short and the initial
light curve rises quickly. In the opposite situation when $^{56}$Ni is
located deeply inside the ejecta, then the dark period can be as long
as a couple days and the initial light curve rises relatively
slowly. The fast initial rise of iPTF16abc and the negligible length
of the dark period both support that $^{56}$Ni is sufficiently mixed
in the ejecta.

>>>>

\title{Optifcal Observations of an Extraordinarily Young Type Ia Supernova iPTF16abc}

@@

More commonly, the only observed light curve of a SN Ia is purely
powered by the radioactive decay of synthesized $^{56}$Ni. Since
$^{56}$Ni atoms are synthesized, mixed and deposited into different
layers inside the ejecta, the SN may experience a dark period after
the SN shock breakout and before the radioactive energy diffuses to
the photosphere \citep{2014ApJ...784...85P}. In the case of strong
mixing, $^{56}$Ni energy can reach the photosphere rapidly. As such,
the dark period is very short and the light curve shows a fast initial
rise. In the opposite case, it may take up to a couple of days for
the $^{56}$Ni energy to diffuse to the photosphere. The initial rise
of the light curve is then moderate \citep{2016ApJ...826...96P}. To
summarize, the initial rise of the light curve of a SN Ia conveys
information on distribution of synthesized $^{56}$Ni.

In this paper, we report observations of an extraordinarily young SN
Ia iPTF16abc, which was discovered by the intermediate Palomar
Transient Factory on 2016 April $3.36$\footnote{all times in this
  paper are in UTC.} at $\textrm{R.A.}=13^h34^m45.49^s$,
$\textrm{Dec.}=+13^d51^m14.3^s$ (J2000) with a $g$-band magnitude of
$21.31\pm0.27$ \citep{2016PASP..128k4502C,2016ATel.8907....1M}. The
transient is spatially coincident with a tidal tail of the galaxy
NGC\,5221 at 100\,Mpc. No activity was detected at the same location
down to $g=22.1$\,mag on April $2.42$. Our spectroscopic follow-up
compaign classified iPTF16abc as a normal SN Ia
\citep{2016ATel.8909....1C}.

This paper is organized as follows: Section \ref{sec:obs} describes
photometric and spectroscopic observations of iPTF16abc. Section
\ref{sec:usual_staff} establishes that iPTF16abc is a normal SN Ia
in NGC\,5221. Section \ref{sec:first_light} analyzes the early
light curve and spectra.

@@

\ycao{TODO: include LCOGT data in both text and figures.}

@@

After discovery, we utilized the rainbow camera of the SED
Machine (\ycao{REF}) mounted on the 60-inch telescope at Palomar
Observatory (P60) to undertake photometric observations in $g$, $r$
and $i$ filters. The image differencing against the archival SDSS
images and forced PSF photometry on the subtracted images were
performed by the Fremling Automated Pipeline
\citep{2016A&A...593A..68F}. The photometry is also calibrated to the
SDSS catalog.

Las Cambres Observatory Global Network (LCOGT) also carry out photometric
observations in the \textit{BVgri} filters with its 1-m telescope network.
\ycao{Some info about data reduction}

@@

\begin{figure*}[!htb]
  \centering
  \includegraphics[width=0.9\textwidth]{spec_sequence.pdf}
  \caption{Low-resolution spectra of iPTF16abc are shown in the
    chronical sequence. The phases in units of days are labeled next
    to corresponding spectra. Telluric absorption bands are grayed
    out.}
  \label{fig:spec_seq}
\end{figure*}

@@

The foreground Galactic extinctioin along the directionof iPTF16abc
has $E(B-V)=0.0279\,\textrm{mag}$ \citep{2011ApJ...737..103S}.

In the highest-resolution spectrum of iPTF16abc by UVES, individual
components of both \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D doublets show
double-absorption profiles (Figure \ref{fig:narrow_features}),
indicating two sources of absorption along the line of sight. Fitting
two Gaussian kernels to each line of the \ion{Na}{1} doublet
simultaneously leads to redshifts of $0.02313820\pm0.00000032$ and
$0.02322408\pm0.00000033$. The total equivalent widths of the
\ion{Na}{1}\,D1 and D2 lines are $0.595\pm0.009\,\textrm{\AA}$ and
$0.609\pm0.008\,\textrm{\AA}$m, respectively. Using the empirical
relation between the equivalent width of \ion{Na}{1}\,D lines and
reddening $E(B-V)$ \citep{2012MNRAS.426.1465P}, we derive
$E(B-V)=0.361\pm0.025\textrm{mag}$.

In the X-shooter spectra of iPTF16abc, we also identify narrow
absorption features \ion{K}{1}\,7665\,\AA and 7699\,\AA at consistent
redshifts. However, the resolution of the X-shooter spectra is not
fine enough to resolve the double-absorption profiles of \ion{Ca}{2}
and \ion{Na}{1}. The X-shooter spectra do not show the diffusive
interstellar band at 5780\,\AA as well.  Albeit some debates (e.g.,
\citealt{2013ApJ...779...38P}), the $E(B-V)$ value derived from
\ion{Na}{1}\,D absorption provides the best estimate in the case of
iPTF16abc.

@@

The \ion{Na}{1}\,D doublet are seen in multiple spectra spanning
from pre-peak to post-peak phases. Despite the instrumental widening of
different instrument configurations, we do not detect obvious variation
in the profiles of the doublet.

@@

We run Supernova Identification (SNID; \citealt{2007ApJ...666.1024B})
on the low-resolution spectrum of iPTF16abc at $+18.8$ and found best
matches to normal SNe Ia. In fact, the characteristic features of a SN
Ia, such as \ion{Si}{2}, \ion{S}{2}, can be easily identified in the
spectra of iPTF16abc (Figure \ref{fig:spec_seq}).

Next we standardize the light curve of iPTF16abc by feeding its P48
and P60 light curves into the \texttt{sncosmo} module\footnote{The
  \texttt{sncosmo} Python module is available at
  \url{https://sncosmo.readthedocs.io/en/v1.4.x/}.} and fit a light
curve model, which composes of the SALT2 template
\citep{2007A&A...466...11G} modified by the line-of-sight extinction
curve \citep{1999PASP..111...63F} with $E(B-V)$ values from Section
\ref{sec:reddening} and $R_V=3.1$. We obtain the rest-frame
\textit{B}-band peak time $\textrm{MJD}_{max}=57499.65\pm0.02$, the coefficient
of the zeroth principle component $x_0=0.0275\pm0.0002$, the
coefficient of the first principle component $x_1=1.200\pm0.043$, and
the color term $c=-0.3353\pm0.0054$. The best-fit model also gives an
unreddened apparent peak magnitude of $m^*_{B}=14.4\,\textrm{mag}$ in
the SN rest frame.

For convenience, in the following sections, we define the best-fit
value $\textrm{MJD}_{max}=57499.65$ as phase $t=0$.

@@

After establishing iPTF16abc as a normal SN ia, we use the latest
calibration \citep{2014A&A...568A..22B} of the Phillips relation
\citep{1993ApJ...413L.105P} using $m^*_{B}$, $x_1$ and $c$ to derive a
distance modulus $\mu=34.66\pm0.03\,\textrm{mag}$ to the SN, provided
that the host galaxy of iPTF16abc has a stellar mass less than
$10^{10}\sm$.

The location of iPTF16abc is spatially coincident with a tidal tail of
galaxy NGC\,5221. \citet{2007A&A...465...71T} derived a distance
modulus of $35.0\pm0.4\,\textrm{mag}$ from the Tully-Fisher
relation. This distance modulus is consistent with that of iPTF16abc.

Separately, \citet{1998A&AS..130..333T} observed the 21-cm line in
this galaxy and measured a redshift of $0.0233303\pm0.000027$.  The
two components in the \ion{Na}{1}\,D have a relative velocity of
$-57.6\pm8.1\,\textrm{km}\,\textrm{s}^{-1}$ and
$-31.8\pm8.1\,\textrm{km}\,\textrm{s}^{-1}$, suggesting that both
absorption resources are probaly located on the tidal tail of
NGC\,5221.

\section{First Light And Explosion Time}
\label{sec:first_light}

In this section, we estimate the first light and explosion time of
iPTF16abc.

@@

Extrapolation of the earliest-phase light curve with a simple model is
usually used to estimate the first light time. In theory,
\citet{1982ApJ...253..785A} derived a quadratic law for an expanding
fireball with a constant temperature. In order to account for the
temperature change, here we utilize a power-law model
\begin{equation}
  \label{eq:broken_power_law}
  f(t) \left\{
    \begin{array}{ll}
      = 0,\ \textrm{when}\ t<t_0 \\
      \propto (t-t_0)^{\alpha},\ \textrm{when}\ t>t_0
    \end{array}
  \right.\ .
\end{equation}
Since the first few detections of the SN were made in the \textit{g}
band, we perform the fit only to the \textit{g}-band light curve. We
experiment the fitting procedure with different time windows and find
that the SN flux between $t=-19\,\textrm{days}$ and
$t=-15\,\textrm{days}$ rises approximately linearly. The joint
distribution of $\alpha$ and $t_0$ for this time window is given in
the left panel of Figure \ref{fig:early_lc_fit} and the best fit model
with $\alpha = 0.94$ and $t_0=-18.47$ is shown in the right
panel. With this best-fit value of $t_0$, our first observation was
made only $0.18\,\textrm{day}$ after the first light of the SN. The
total rise time to the \textit{B}-band peak is $18.47$ days.

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.45\textwidth]{rise_time_power_law_index.pdf}
  \includegraphics[width=0.45\textwidth]{early_lc.pdf}
  \caption{Broken Power low fitting to the early $g$-band light
    curve.
    \textit{Left:} the joint distribution of $t_0$ and
    $\alpha$. The cross marker denotes the best-fit parameters. The
    inner and outer contours represent the $68\%$ and $99.7\%$
    confidence levels.
    \textit{Right:} the best-fit model of $\alpha=0.94$ and
    $t_0=-18.47\,\textrm{days}$ is illustrated against the
    data.
  }
  \label{fig:early_lc_fit}
\end{figure*}

Since $t=-15\,\textrm{days}$, the \textit{g}-band light curve rises
significantly faster than the best-fit model above, indicating a
greater value of the power-law index $\alpha$. In fact, the
light curve between $t=-14$ and $t=-8$\,days can be fitted with a
power law of index $1.40$.

A more sophisticated broken power-law model has also been used to fit
SN early-phase light curves (e.g.,
\citealt{2013ApJ...778L..15Z,2014ApJ...783L..24Z,
  2016arXiv161202097Z,2016arXiv161202097Z}). Our above result has
indicates that the \textit{g}-band light curve of iPTF16abc can also
be fitted by a broken-power law with a power index changing from
$0.94$ to $1.40$. In fact, many young SNe, including SN2011fe
\cite{2016arXiv161202097Z} and iPTF16abc, have early light curves that
follow broken power-laws \citep{2016arXiv161202725Z}). Our above
result has indicates that the \textit{g}-band light curve of iPTF16abc
can also be fitted by a broken-power law with a power index changing
from $0.94$ to $1.40$.

The initial rise of a SN depends on the distribution of $^{56}$Ni
in the ejecta. Comparing the rapid initial rise of iPTF16abc to
theoretical calculations in \citet{2016ApJ...826...96P} suggests
strong mixing of $^{56}$Ni in the iPTF16abc ejecta. The radioactive
energy from $^{56}$Ni can quickly diffuse to the SN photosphere and
power its light curve after the SN explosion.

@@

The first light of a SN is not necessarily its explotion time
$t_{exp}$.  Thanks to different deposition of $^{56}$Ni in the SN
ejecta, as pointed out in \citet{2014ApJ...784...85P}, the SN may
experience a dark period before the radioactive energy from the
shallowest $^{56}$Ni layer escapes the ejecta. Therefore, 
\citet{2014ApJ...784...85P} suggests to examine evolution of absorption
line velocities, because the velocities do not depend on the $^{56}$Ni
mixing and are expected to evolve roughly as $v\propto(t-t_{exp})^{-0.22}$.

We focus on measuring velocities of the strongest \ion{Si}{2}\,6355
line, as the second most commonly used \ion{Ca}{2} IR triplet is very
weak in iPTF16abc. Visual inspection of the early-phase spectra shows
no sign of any high-velocity component of the \ion{Si}{2} line and
that the red wing of \ion{Si}{2} overlaps the \ion{C}{2}\,6580
line. Therefore, we fit simultaneously two gaussian kernels for the
two lines and a linear term to account for the continuum component.
Then the velocities are calculated at the minimum of the \ion{Si}{2}
gaussian kernel.

\begin{figure*}[!thb]
  \centering
  \includegraphics[width=0.45\textwidth]{exp_date_chi2.pdf}
  \includegraphics[width=0.45\textwidth]{SiIIVelocity.pdf}
  \caption{Constraints on $t_{exp}$ from fitting the velocity
    evolution of \ion{Si}{2}. \textit{Left panel:} the dashed, solid
    and dash-dotted curves show $\chi^2$ for fitting power laws with
    indices $-0.20$, $-0.22$ and $-0.24$, respectively. The red
    vertical line and the orange region indicate $t_0$ and its
    3-$\sigma$ confidence interval from Section
    \ref{sec:lc_fit}. \textit{Right panel:} Observed \ion{Si}{2}\,6355
    velocities with the best-fit power-law velocity with an index of
    $-0.22$.}
  \label{fig:velocity_t_exp}
\end{figure*}

Fitting the measured velocities of \ion{Si}{2}\,6355 line to the
$v\propto(t-t_{exp})^{-0.22}$ model leads to the best-fit explosion
time $t_{exp}=-17.95\,\textrm{days}$ with a 3-$\sigma$ confidence
interval between $-17.4\,\textrm{days}$ and $-18.3\,\textrm{days}$
(Figure \ref{fig:velocity_t_exp}). We further test the robustness of
a fixed power-law index by altering the index to $-0.20$ and $-0.24$
and find consistent results within respective 3-$\sigma$ confidence
intervals.

The derived $t_{exp}$ is roughly consistent to $t_0$ estimated in
Section \ref{sec:lc_fit}, confirming that the dark period of iPTF16abc
is very short. This conclusion is consistent with the inference from the
rapid initial rise of iPTF16abc in Section \ref{sec:lc_fit}.

Among the several SNe with measurements of $t_{exp}$ and $t_0$
\citep{2014ApJ...784...85P,2016ApJ...826..144S}, iPTF16abc has the
shortest dark period and longest rise time ($18.47$ days). In
comparison, SN2011fe is inferred to have a dark period of $\sim 1$
day \citep{2014ApJ...784...85P} and a rise time of $17.7$ days
\citep{2013A&A...554A..27P}; ASASSN-14lp has a dark period of a couple
of days and a rise time of $16.94$ days. This seems to suggest that
the time difference between SN explosion and the \textit{B}-band maximum,
which equals to summation of the dark period and the rise time, is a
constant for a SN Ia. This is perhaps because the \textit{B}-band maximum
time only depends on the total amount of $^{56}$Ni.

\subsection{Strong and Short-Lived Carbon Features}
\label{sec:carbon}

Interestingly, the first few spectra of iPTF16abc show clear detection
of \ion{C}{2}\,6580 and \ion{C}{2}\,7234 lines. Their velocity
evolution is shown in Figure \ref{fig:velocity_t_exp} as well.

The \ion{C}{2} lines only appear in the very early phases. The
equivalent widths of detected \ion{C}{2}\,6580 and \ion{C}{2}\,7234
lines are shown in Figure \ref{fig:ew}, and are compared to that of
the \ion{Si}{2}\,6355 line. Especially in the first spectrum taken at
$t=-15.8$ days, the \ion{C}{2} equivalent widths are stronger than
that of \ion{Si}{2}.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.45\textwidth]{EW.pdf}
  \caption{Equivalent Width of \ion{Si}{2}\,6355, \ion{C}{2}\,6580 and
    \ion{C}{2}\,7234 lines}
  \label{fig:ew}
\end{figure}

The strong and short-lived \ion{C}{2} features suggest that
the outermost layers of the ejecta contain a certain amount of
carbon which is responsible for the \ion{C}{2} absorption. The
abundance of carbon drops dramatically into the inner layers
of the ejecta.

Carbon signatures are seen in over $1/4$ of all normal SNe Ia before
maxima
\citep{2011ApJ...732...30P,2012MNRAS.425.1917S,2011ApJ...743...27T},
but the signatures are usually weak. Even in SN2011fe, the \ion{C}{2}
features are not strong in the first spectra
\citep{2012ApJ...752L..26P}.  The only normal SN Ia known to have
strong \ion{C}{2} features at very early phases is SN2013dy
\citep{2013ApJ...778L..15Z}. Unlike iPTF16abc, the equivalent widths
of \ion{C}{2} features in SN2013dy are as large as that of
\ion{Si}{2}\,6355.

\subsection{Discussions and Physical Constraints}
\label{sec:lc_energy}

The early radiation of a SN Ia may have multiple resources: SN
shockbreakout, SN-companion collision, and radioactive activity. Since
each provides interesting constraints on the progenitor properties, we
explore all the three posibilities.

\subsubsection{SN Shock Breakout}

The shock breakout of a SN Ia lasts for a fraction of a second due to
the small size of the exploding star. However, the subsequent cooling
phase may last longer (e.g., \citealt{2010ApJ...708..598P}).
Following the analysis of SN2011fe in \citet{2012ApJ...744L..17B}, we
compare the early-phase \textit{g}-band light curve of iPTF16abc with
two cooling models \citep{2011ApJ...728...63R, 2010ApJ...708..598P}
and reach a not very constraining conclusion that the radius of the
progenitor star of iPTF16abc should be $<1\sr$. In fact, the cooling
emission is negligible at the time of the first detection of
iPTF16abc.

\subsubsection{SN companion}

If iPTF16abc is born in a single-degenerate channel, with an odd of
$\lesssim10\%$, we expect to see emission produced by SN ejecta
slamming into the companion. However, according to calculations by
\citet{2010ApJ...708.1025K}, collision between SN ejecta and the
companion star produces thermal emission with a spectrum that peaks
in the ultraviolet, and the Rayleigh-Jeans tail emission in the \textit{g}
band is very weak.

In the case of SN2012cg, in order to illustrate a blue excess in the
early light curve, \citet{2016ApJ...820...92M} fitted a power-law
model to its light curves between $t=-14$ days and $t=-8$ days and
subtracted the extrapolated fluxes at early epochs. Then these authors
attribute the excess to SN-companion collision. In our iPTF16abc, we
follow the same procedure and also find a similar excess
(Figure \ref{fig:against_sn2012cg}).

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.45\textwidth]{another_early_lc.pdf}
  \caption{The early-phase \textit{g}-band light curve of iPTF16abc is
    compared against the best-fit power-law model to the data segment
    between $t=-14$ and $t=-8$ days. Note the ``excess'' between the
    data and the model during $t=-18$ and $t=-15$ days}
  \label{fig:against_sn2012cg}
\end{figure}

Actually, for any early light curve that can be approximated by a
broken power law, as long as the asymptotic power-law index near $t=0$
is less than that at late time, one can always find such an excess
near $t=0$. There are not many SNe that were observed at
extraordinarily early phases. The odd that both SN2012cg and iPTF16abc
show SN-companion collision signautres is very small.

Separately, our first \textit{Swift} observation of iPTF16abc was made
about two days after explosion. However, the ultraviolet light curve of
iPTF16abc does not show any excess.

\subsubsection{Explosion Mechanism}

Our extraordinarily early observations of iPTF16abc put two constraints
on the mechanism of the SN explosion: first, during the SN explosion,
synthesized $^{56}$Ni should be strongly mixed in the ejecta; second,
unburned carbon atoms are accelerated to very high velocities in the ejecta.
In addition, the commonly seen \ion{Ca}{2} IR triplet is weak in iPTF16abc,
implying a little amount of calsium in the ejecta.

\section{Conclusion}
\label{sec:conclusion}

In this paper, we present observations of a normal Type Ia supernova
iPTF16abc, which is located at one tidal tail of NGC\,5221. Our extraordinarily
early-phase optical observations of this supernova show that:
\begin{itemize}
\item Extrapolation of the early light curve shows that our first observation
  was made only $0.18$ days after the first light of the SN.
\item The rapid initial rise and analysis of the line velocity both shows that
  the dark period of iPTF16abc is very short. Hence, the first light time of
  the SN is a good approximation to the time of the SN explosion.
\item Strong and short-lived carbon features in the early spectra of iPTF16abc
  indicates concentration of unburned carbon in the fastest-moving part of the
  ejecta.
\end{itemize}

\acknowledgements
